Cassini State Capture

### 2019-10-17 writeups/Su_disk_disp.tex
Modified a ton more plots

% TODO
% figure out whether can draw vertical lines on 3testo plots for regimes
% run simulations up to III -> III (not today)

@@

Denote $\hat{s}$ spin of planet, $\hat{l}$ angular momentum of planet, and
$\hat{l}_d$ angular momentum of the surrounding disc. We approximate $S \ll L
\ll L_d$, so $\hat{l}_d$ is approximately constant and $\hat{l}$ experiences
negligible backreaction torque from $\hat{s}$. Much of this treatment borrows
from \citealp{anderson2018teeter} and \citealp{millholland_disk}. We consider
precession equations
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}}\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d}\label{eq:dldt}.
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, $M_\star$ is
the mass of the host star, and $M_p, R_p, a_p, \Omega_p$ are the mass, radius,
semi-major axis and angular spin frequency of the planet respectively.
\autoref{eq:wld} is written for a disk of mass $M_d$ located at characteristic
distance $r_d$ from the host star \citep[see][for a power-law disk
profile]{millholland_disk}, $n \equiv \sqrt{GM_\star/a_p^3}$ is the planet's
orbital mean motion. The three angles formed by the three vectors are notated
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

Let us transform to frame rotating about $\hat{l}_p$ with frequency $g \equiv
-\omega_{ld}\cos I < 0$, where $\hat{l}$ fixed. In this frame, $\hat{l}_p$
also remains fixed and $\hat{s}$ evolves as:
\begin{equation}
    \rd{\hat{s}}{t} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_d},
\end{equation}
where $\alpha \equiv \omega_{sl} > 0$ follows standard notation. By defining
$\eta$ and nondimensionalized time $\tau$ as
\begin{align}
    \eta &\equiv \frac{\abs{g}}{\alpha}\label{eq:eta},\\
    \tau &\equiv \alpha t,
\end{align}
we can non-dimensionalize the equation of motion as
\begin{equation}
    \rd{\hat{s}}{\tau} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_d}. \label{eq:dsdt_base}
\end{equation}

Following~\cite{millholland_disk}, we allow $g$ to vary in time owing to a
disk with decaying mass
\begin{equation}
    M_d(t) = M_d(0)e^{-t/t_d}.
\end{equation}
As $\alpha$ is held constant, this equates to $\eta$ decaying as $\rd{\eta}{t} =
-\frac{\eta}{t_d}$ for some decay time $t_\eta$. We nondimensionalize by
defining $\epsilon \equiv \frac{1}{\alpha t_\eta}$, so
\begin{equation}
    \rd{\eta}{\tau} = -\epsilon \eta.\label{eq:deta_dt}
\end{equation}
\autoref{eq:dsdt_base} and \autoref{eq:deta_dt} together constitute our system
of study.

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CSs). There are either two or four solutions, depending on the values
of $\eta, I$. Following standard convention and nomenclature (see
\autoref{fig:cs_locs}), CSs 1, 3, 4 have $\phi = 0, \theta < 0$ and CS2 has
$\phi = \pi, \theta > 0$. Of these, CSs 1, 2, 3 are stable while CS4 is unstable
(this follows from linearization of \autoref{se:H_eom}). When $\eta < \eta_c$,
where
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_c$, only CSs 2, 3 exist
\citep{henrard1987,ward2004I}. The CS obliquities as a function of $\eta$ are
shown in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{$\theta$ values of the Cassini states over time, following the same
    coloring scheme as \autoref{fig:eq_1contours}. Note that $\theta \in [-\pi,
    \pi]$ is the traditional definition of the polar angle
    \citep[see e.g.][]{colombo1966,peale1969,henrard1987}.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_c$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. The separatrix, three
zones, and their relations to the CSs are shown in \autoref{fig:eq_1contours}.
Note that trajectories in zone II librate about CS2 while those in zones I/III
circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$, where warmer
    colors denote more positive values. The black solid line is the separatrix,
    which only exists for $\eta < \eta_c$. The three zones, divided by the
    separatrix, are labeled. The interior of the separatrix, shaded in grey, is
    formally only defined for $\eta < \eta_c$, but we may identify the phase
    space trajectories that flow into zone II when evolved forward in time; this
    is the shaded region in the top left plot, bounded by the dotted black
    line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{align*}
    z_0 &= \eta\cos I, &
    \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
    \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
        \chi^2 \cos^2\theta_4 + 1},&
    T &= 2\chi \frac{\cos \theta_4}{
        \chi^2 \cos^2\theta_4 - 1},\\
\end{align*}
the areas for $\eta < \eta_c$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_I &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_c$, but a natural extension
exists: evolve an initial point $p$ under adiabatic decrease of $\eta > \eta_c$
until the separatrix appears at $\eta = \eta_c$, then identify $p$ with the zone
it is in at $\eta_c$. Since phase space area is conserved under adiabatic
evolution, this implies $A_i\p{\eta > \eta_c} = A_i(\eta_c)$. This extension is
also reflected in the dashed boundary in \autoref{fig:eq_1contours}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional $A_{i}(\eta)$ as given by
    \autoref{se:area_ward}. Dotted lines correspond to small $\eta$
    approximations discussed in \autoref{ss:app_transition}. Dashed lines denote
    effective values of $A_{II}, A_{III}$ for $\eta > \eta_c$, denoting the
    area of points that would flow into either area under adiabatic decrease of
    $\eta$ to below $\eta_c$.}\label{fig:eq_areas}
\end{figure}

@@

One criterion for adiabaticity is that the separatrix crossing timescale be
longer than the the resonant libration period \citep{ward2004I}. The resonant
libration period is just the libration period about CS2, which can be found by
linearizing \autoref{se:H_eom} about CS2. This results in constraint
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}}_{cross} = \epsilon &\lesssim
            \frac{1}{T_{lib}},\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta_\star\sin\theta \sin I
            \p{1 + \eta_\star \frac{\sin I}{\sin^3\theta}}}.\label{eq:ad_constr}
\end{align}
This formula differs from that presented in \citealt{millholland_disk}.

\subsection{Dynamical Outcomes}\label{ss:ad_ensemble}

We consider the evolution of a system with arbitrary initial $\theta_{sd, i}
\equiv \theta_{sd}\p{t = 0}$ and $\eta_i \gg 1$. We are interested in the
distribution of final $\theta_f$ values after $\eta$ decreases to $\eta_f \ll 1$
(after the disk has dissipated to negligible mass). While it is likely that
$\theta_{sd, i} \approx 0$ as the planet forms in the disk, we explore all
$\theta_{sd, i} \in [0, \pi]$ for completeness.

We evenly sample $101$ values of $\theta_{sd, i}$, and for each $\theta_{sd, i}$
value, we pick $101$ $\hat{s}$ initial conditions on the ring fixed by
$\theta_{sd, i}$. To be concrete, we choose $\eta_i = 10\eta_c$. For each
initial condition, we evolve Equations~\ref{eq:dsdt_base} and~\ref{eq:deta_dt}
until $\eta \leq 10^{-5}$, where $\hat{s}$ precesses about $\hat{l}$ with
uniform final obliquity $\theta_f$. This transfer function
$\theta_f\p{\theta_{sd, i}}$ is our primary result, and is shown for $I =
5^\circ, 10^\circ, 20^\circ$ in \autoref{fig:ad_ensemble},
\autoref{fig:3_ensemble_10_35}, and \autoref{fig:3_ensemble_20_35} respectively.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{f}\p{\theta_{sd, i}}$, overlaid with semi-analytic
    predictions of the $\theta_{ f}$ for each of the four nontrivial
    dynamical tracks in colored lines. Bottom: Semi-analytic probabilities of
    each of the dynamical tracks for each $\theta_{sd, i}$. The five regimes of
    $\theta_{sd, i}$ values correspond to the five regimes of $A_i$ discussed in
    \autoref{ss:zone_transitions} when using \autoref{eq:ai_qsd} to relate
    $A_i\p{\theta_{sd, i}}$. }\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}

\subsection{Separatrix Crossing and Evolutionary Tracks
}\label{ss:zone_transitions}

In the limit $\eta \to 0$, zone II disappears and all trajectories circulate at
constant final $\theta$ values, notated $\theta_f$. We run simulations that
terminate at $\eta = 10^{-{-3.5}}$ and measure $\theta_{f}$. We vary the initial
value of $\theta_{sd}$, notated $\theta_{sd, i}$, among our simulations.

In the absence of separatrix encounters, the enclosed phase space area $A \equiv
\oint \cos\theta \;\mathrm{d}\phi$ is an adiabatic invariant. Here, we adopt
convention where $A$ is a \emph{signed} area
\begin{equation}
    A = \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.
\end{equation}
This has the advantage of (i) having a removable singularity when the trajectory
crosses $\cos \theta = 1$, and (ii) being easily expressible as combinations of
the $A_i$ of \autoref{se:area_ward}. The bounds of the integral are either a
libration ($\phi$ returns to its original value with the same $\dot{q\phi}$
sign) or circulation cycle ($\phi$ advances to $\phi \pm 2\pi$). Note lastly
that $A$ has simple physical interpretation of are enclosed on the unit sphere
by $\hat{s}$ over one cycle measured relative to $+\hat{z}$:
\begin{equation*}
    \int \int\limits_{\cos \theta(\phi)}^1
        \mathrm{d}\cos\theta\;\mathrm{d}\phi = \int \p{1 - \cos \theta(\phi)}
            \;\mathrm{d}\phi.
\end{equation*}

Observe that when $\eta \gg 1$, trajectories librate about $\hat{l}_d$, or
equivalently CS2, with approximately constant $\theta_{sd}$, meaning they
enclose phase space area
\begin{equation}
    A_i = 2\pi\p{1 - \cos \theta_{sd, i}}.\label{eq:ai_qsd}
\end{equation}
In our simulations, we start with finite $\eta$, so we use the angular distance
to CS2 instead of $\theta_{sd, i}$ above, as the trajectory initially librates
about CS2 and not $\hat{l}_p$.

When a trajectory encounters the separatrix, $A$ will change discontinuously as
the trajectory moves between zones, but its change is easily understood
\citep{henrard1982}. We can classify trajectories by the separatrix encounters
they undergo. Initially, in the $\eta > \eta_c$ regime, only zones $II, III$
exist. Conversely, at the end of the simulation when $\eta \to 0$, only zones
$I, III$ exist. Most simply, one might expect four sequences of transitions
between zones (we call these dynamical ``tracks'') to manifest: from one of
$\z{I, III}$ to one of $\z{II, III}$. In addition to these four tracks, a
fifth track $III \to II \to I$ is observed.

Beyond simply identifying the possible tracks, we may clearly delineate the
initial conditions for which each track can be encountered. Separatrix crossing
is a probabilistic process whose outcomes are well understood by
\citealp{henrard1982,henrard1987}. Their resultss may be simply summarized as
follows: if a zone $i$ is shrinking and zone $j$ is expanding, the probability
of transition from zone $i$ to zone $j$ is given
\begin{equation}
    \Pr_{i \Rightarrow j} = -\frac{\pd{A_j}{t}}{ \pd{A_i}{t}}
        = -\frac{\pd{A_j}{\eta}}{ \pd{A_i}{\eta}}\label{eq:henrard_hop}
\end{equation}
This can be used directly in conjunction with \autoref{se:area_ward} to
understand for what initial conditions each track can be observed and with what
probabilities.

Below, we will describe the evolution of $A$ throughout each of the five tracks,
as well as determine the initial conditions and probabilities associated with
each track:
\begin{itemize}
    \item $II \to I$ --- This track is depicted in \autoref{fig:ad_21}.
        $\hat{s}$ starts librating about CS2 in zone II, enclosing some initial
        phase space area $A_i$. The trajectory is able to librate without
        separatrix encounter until $A_{II}(\eta_\star) = A_i$. At this point, as
        the trajectory transitions to a circulating trajectory in zone I
        immediately bordering the separatrix, it will encompass
        $-A_I(\eta_\star)$ phase space area. Conservation of phase space area
        then predicts that as $\eta \to 0$, the final enclosed phase space area
        will still be $A_f = -A_I$, and the final obliquity $\theta_f$ must be
        given by
        \begin{equation}
            2\pi\p{1 - \cos \theta_f} = A_f. \label{eq:qfaf}
        \end{equation}

        This transition can only occur when the initial condition begins in zone
        II, requiring $A_i < A_{II}(\eta_c)$, where $A_{II}(\eta_c)$ is known in
        closed form \citep{ward2004I}
        \begin{equation}
            A_{II}\p{\eta_c} = 4\pi \s{
                1 - \p{1 + \tan^{2/3}I}^{-3/2}}.
        \end{equation}
        Then, at some known $\eta_\star$ satisfying $A_{II}(\eta_\star) = A_i$,
        the trajectory is ejected from the separatrix and enters zone I with
        probability
        \begin{equation}
            \Pr_{II \to I} = -\at{\frac{\pd{A_I}{\eta}}{\pd{A_{II}}{\eta}}}
                _{\eta_\star}.
        \end{equation}
        Note that since $\pd{A_i}{\eta} < 0$ everywhere, while
        $\pd{A_{II}}{\eta} > 0$ at all possible $\eta_\star$ for an initial
        condition starting in zone II, this track always has nonzero
        probability.

>>>>

@@

Denote $\hat{s}$ spin of planet, $\hat{l}$ angular momentum of planet, and
$\hat{l}_d$ angular momentum of the surrounding disc. We approximate $S \ll L
\ll L_d$, so $\hat{l}_d$ is approximately constant and $\hat{l}$ experiences
negligible backreaction torque from $\hat{s}$. Much of this treatment borrows
from \citealp{anderson2018teeter} and \citealp{millholland_disk}. We consider
precession equations
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}}\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d}\label{eq:dldt}.
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
% = 1/(152600 yr) (kp = 6kqp, Mstar = msun, mp = mJ, Rp = RJ, ap = AU, Wp = 2pi
% / (10 days))
% = 1 / (573362 yr) (Md/Mstar = 0.01, ap/rd = 1/30)
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, $M_\star$ is
the mass of the host star, and $M_p, R_p, a_p, \Omega_p$ are the mass, radius,
semi-major axis and angular spin frequency of the planet respectively.
\autoref{eq:wld} is written for a disk of mass $M_d$ located at characteristic
distance $r_d$ from the host star \citep[see][for a power-law disk
profile]{millholland_disk}, $n \equiv \sqrt{GM_\star/a_p^3}$ is the planet's
orbital mean motion. The three angles formed by the three vectors are notated
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

Let us transform to frame rotating about $\hat{l}_p$ with frequency $g \equiv
-\omega_{ld}\cos I < 0$, where $\hat{l}$ fixed. In this frame, $\hat{l}_p$
also remains fixed and $\hat{s}$ evolves as:
\begin{equation}
    \rd{\hat{s}}{t} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_d},
\end{equation}
where $\alpha \equiv \omega_{sl} > 0$ follows standard notation. By defining
$\eta$ and nondimensionalized time $\tau$ as
\begin{align}
    \eta \equiv{}& \frac{\abs{g}}{\alpha}\label{eq:eta},\\
        ={}& 3.75 \p{\frac{k_p}{6k_{qp}}}
            \p{\frac{m_p}{0.001 M_\star}}
            \p{\frac{a_p}{1\;\mathrm{AU}}\frac{R_J}{R_p}}^3
            \p{\frac{2\pi/\p{10\;\mathrm{days}}}{\Omega_p}}\nonumber\\
        &\p{\frac{M_d}{0.01 M_{\star}}}
            \p{\frac{30 a_p}{r_d}}^3
            \p{\frac{n}{2\pi / \p{1\;\mathrm{yr}}}}
            \p{\frac{\cos I}{\cos 5^\circ}},\nonumber\\
    \tau \equiv{}& \alpha t,
\end{align}
we can non-dimensionalize the equation of motion as
\begin{equation}
    \rd{\hat{s}}{\tau} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_d}. \label{eq:dsdt_base}
\end{equation}

Following~\cite{millholland_disk}, we allow $g$ to vary in time owing to a
disk with decaying mass
\begin{equation}
    M_d(t) = M_d(0)e^{-t/t_d}.
\end{equation}
As $\alpha$ is held constant, this equates to $\eta$ decaying as $\rd{\eta}{t} =
-\frac{\eta}{t_d}$ as well. We nondimensionalize:
\begin{align}
    \epsilon \equiv{}& \frac{1}{\alpha t_d},\\
        ={}& 0.015 \p{\frac{k_p}{6k_{qp}}}
            \p{\frac{m_p}{0.001 M_\star}}
            \p{\frac{a_p}{1\;\mathrm{AU}}\frac{R_J}{R_p}}^3\nonumber\\
        &\p{\frac{2\pi/\p{10\;\mathrm{days}}}{\Omega_p}}
            \p{\frac{10\;\mathrm{Myr}}{t_d}},\\
    \rd{\eta}{\tau} ={}& -\epsilon \eta.\label{eq:deta_dt}
\end{align}
\autoref{eq:dsdt_base} and \autoref{eq:deta_dt} together constitute our system
of study.

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CSs). There are either two or four solutions, depending on the values
of $\eta, I$. Following standard convention and nomenclature (see
\autoref{fig:cs_locs}), CSs 1, 3, 4 have $\phi = 0, \theta < 0$ and CS2 has
$\phi = \pi, \theta > 0$. Of these, CSs 1, 2, 3 are stable while CS4 is unstable
(this follows from linearization of \autoref{se:H_eom}). When $\eta < \eta_c$,
where
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_c$, only CSs 2, 3 exist
\citep{henrard1987,ward2004I}. The CS obliquities as a function of $\eta$ are
shown in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$. Note that
    $\theta \in [-\pi, \pi]$ is the traditional definition of the polar angle
    \citep[see e.g.][]{colombo1966,peale1969,henrard1987}.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_c$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. The separatrix, three
zones, and their relations to the CSs are shown in \autoref{fig:eq_1contours}.
Note that trajectories in zone $II$ librate about CS2 while those in zones
$I$/$III$ circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$, where warmer
    colors denote more positive values. The black solid line is the separatrix,
    which only exists for $\eta < \eta_c$. The three zones, divided by the
    separatrix, are labeled. The Cassini states are labeled and have the same
    colors as \autoref{fig:cs_locs}. The interior of the separatrix, shaded in
    grey, is formally only defined for $\eta < \eta_c$, but we may identify the
    phase space trajectories that flow into zone $II$ when evolved forward in
    time; this is the shaded region in the top left plot, bounded by the dotted
    black line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{align*}
    z_0 &= \eta\cos I, &
    \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
    \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
        \chi^2 \cos^2\theta_4 + 1},&
    T &= 2\chi \frac{\cos \theta_4}{
        \chi^2 \cos^2\theta_4 - 1},\\
\end{align*}
the areas for $\eta < \eta_c$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_I &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_c$, but a natural extension
exists: evolve an initial point $p$ under adiabatic decrease of $\eta > \eta_c$
until the separatrix appears at $\eta = \eta_c$, then identify $p$ with the zone
it is in at $\eta_c$. Since phase space area is conserved under adiabatic
evolution, this implies $A_i\p{\eta > \eta_c} = A_i(\eta_c)$. This extension is
also reflected in the dashed boundary in \autoref{fig:eq_1contours}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional $A_{i}(\eta)$ as given by
    \autoref{se:area_ward}. Dotted lines correspond to small $\eta$
    approximations discussed in \autoref{s:app_transition}. Dashed lines for
    $\eta > \eta_c$ effective values of $A_{II}, A_{III}$ for $\eta > \eta_c$,
    denoting the area of points that would flow into either area under adiabatic
    decrease of $\eta$ to below $\eta_c$.}\label{fig:eq_areas}
\end{figure}

@@

One criterion for adiabaticity is that the separatrix crossing timescale be
longer than the the resonant libration period \citep{ward2004I}. The resonant
libration period is just the libration period about CS2, which can be found by
linearizing \autoref{se:H_eom} about CS2. This results in constraint
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}}_{cross} = \epsilon &\lesssim
            \frac{1}{T_{lib}},\nonumber\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta_\star\sin\theta \sin I
            \p{1 + \eta_\star \frac{\sin I}{\sin^3\theta}}}.\label{eq:ad_constr}
\end{align}
This formula differs from that presented in \citealt{millholland_disk}.

\subsection{Adiabatic Evolution Outcomes}\label{ss:ad_ensemble}

We consider the evolution of a system with arbitrary initial $\theta_{sd, i}
\equiv \at{\theta_{sd}}_{t = 0}$ and $\eta_i \equiv \at{\eta}_{t = 0} \gg 1$. We
are interested in the distribution of final obliquities after $\eta$ decreases
to $\eta_f \ll 1$ (after the disk has dissipated to negligible mass). While it
is likely that $\theta_{sd, i} \approx 0$ as the planet forms in the disk, we
explore all $\theta_{sd, i} \in [0, \pi]$ for completeness.

We evenly sample $101$ values of $\theta_{sd, i}$, and for each $\theta_{sd, i}$
value, we pick $101$ $\hat{s}$ initial conditions approximately on the ring
fixed by $\theta_{sd, i}$\footnote{For each $\theta_{sd, i}$ we choose an
initial condition on the libration cycle with initial condition $\p{\theta_2 +
\theta_{sd, i}, \phi_2}$ (where $\p{\theta_2, \phi_2}$ are the coordinates of
CS2), such that all initial conditions for a given $\theta_{sd, i}$ enclose the
same initial phase space; see \autoref{sss:a_evo}.}. To be concrete, we choose
$\eta_i = 10\eta_c$. For each initial condition, we evolve
Equations~\ref{eq:dsdt_base} and~\ref{eq:deta_dt} until $\eta = 10^{-5}$. For
such small $\eta$, $\hat{s}$ precesses about $\hat{l}$ with uniform final
obliquity $\theta_f$. This transfer function $\theta_f\p{\theta_{sd, i}}$ is our
primary result, and is shown for $I = 5^\circ, 10^\circ, 20^\circ$ in
\autoref{fig:ad_ensemble}, \autoref{fig:3_ensemble_10_35}, and
\autoref{fig:3_ensemble_20_35} respectively. The blue dots represent the results
of the simulation, which are in excellent agreement with the colored tracks.
These colored tracks are calculated semi-analytically using methods explained in
the following section.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{f}\p{\theta_{sd, i}}$, overlaid with semi-analytic
    predictions of the $\theta_{ f}$ for each of the four nontrivial
    dynamical tracks in colored lines. Black dotted lines correspond to
    small-$\theta_{sd, i}$ predictions derived in \autoref{s:app_transition}.
    Bottom: Semi-analytic probabilities of each of the dynamical tracks for each
    $\theta_{sd, i}$. The five regimes of $\theta_{sd, i}$ values correspond to
    the five regimes of $A_i$ discussed in \autoref{ss:zone_transitions}. Again,
    black dotted lines correspond to simple analytic estimates. Vertical black
    dashed lines correspond to the analytical predictions of where each
    dynamical track is accessible, as described in
    \autoref{ss:zone_transitions}.}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I = 20^\circ$. The $III
    \to II \to I$ track is not visible since the maximum $A_{II}$ is very near
    $\eta_c$ for $I = 20^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}

\subsection{Adiabatic Evolution Theory}\label{ss:zone_transitions}

The evolutionary tracks that govern the behavior of $\theta_f\p{\theta_{sd,
i}}$ correspond to particular sequences of separatrix crossings. Coarsely, they
can be understood using a combination of (i) how the enclosed phase space area
by the trajectory evolves across each separatrix crossing, and (ii) the
associated probabilities with each separatrix crossing.

\subsubsection{Governing Principles: Evolution of $A$}\label{sss:a_evo}

First, we consider how the enclosed phase space area by a trajectory evolves
over time. In the absence of separatrix encounters, the enclosed phase space
area $\oint \cos\theta \;\mathrm{d}\phi$ is an adiabatic invariant. In
particular, we adopt convention where
\begin{equation}
    A \equiv \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.\label{eq:a_oint}
\end{equation}
This definition of $A$ has the advantage of (i) being continuous at $\cos \theta
= 1$, and (ii) being easily expressible as combinations of the $A_i$ of
\autoref{se:area_ward}. The bounds of the integral are either a libration
($\phi$ returns to its original value with the same $\dot{q\phi}$ sign) or
circulation cycle ($\phi$ advances to $\phi \pm 2\pi$). Note lastly that $A$ has
simple physical interpretation of are enclosed on the unit sphere by $\hat{s}$
over one cycle measured relative to $+\hat{z}$:
\begin{equation*}
    \int \int\limits_{\cos \theta(\phi)}^1
        \mathrm{d}\cos\theta\;\mathrm{d}\phi = \int \p{1 - \cos \theta(\phi)}
            \;\mathrm{d}\phi.
\end{equation*}

Observe that when $\eta_i \gg 1$, trajectories librate about $\hat{l}_d$ with
constant $\theta_{sd}$, meaning they enclose initial phase space area
\begin{equation}
    A_i \equiv \at{A}_{t = 0}
        = 2\pi\p{1 - \cos \theta_{sd, i}}.\label{eq:ai_qsd}
\end{equation}
Complications arise when considering finite $\eta_i$, as trajectories near CS2 or
CS3 librate about these equilibria respectively, rather than $\hat{l}_d$, and
\autoref{eq:ai_qsd} is no longer exact. In practice, \autoref{eq:ai_qsd} holds
very well when defining $\theta_{sd, i}$ as the angular distance to CS2; an
exception is discussed in \autoref{sss:evol_traj}.

Beginning at the last separatrix crossing, the final enclosed phase space area
$A_f$ will be conserved for all time. As $\eta \to 0$, trajectories circulate at
constant obliquity $\theta_f$, related to $A_f$ by
\begin{equation}
    2\pi\p{1 - \cos \theta_f} = A_f. \label{eq:qfaf}
\end{equation}

$A$ is not conserved when the trajectory encounters the separatrix. Howover,
its change is easily understood \citep{henrard1982}. In essence, when the
trajectory crosses the separatrix, it continues to evolve adjacent to the
separatrix. So if a separatrix crossing results in a zone I trajectory, the
resultant $A$ can be approximated by integrating \autoref{eq:a_oint} along the
upper leg of the separatrix. Pictorally, this can be seen in the bottom plots of
\autoref{fig:ad_21}.

\subsubsection{Governing Principles: Probabilistic Separatrix Crossing}

When a trajectory experiences separatrix crossing, it transitions into nearby
zones probabilistically. This process is well understood by
\citealp{henrard1982,henrard1987}. Their results may be simply summarized as
follows: if a zone $i$ is shrinking while adjacent zones $j, k$ are expanding
such that the sum of their areas is constant, the probability of transition from
zone $i$ to zone $j$ is given
\begin{align}
    \Pr\p{i \to j} = -\frac{\pd{A_j}{t}}{ \pd{A_i}{t}}
        = -\frac{\pd{A_j}{\eta}}{ \pd{A_i}{\eta}},\\
    \Pr\p{i \to k}
        = -\frac{\pd{A_k}{\eta}}{ \pd{A_i}{\eta}}.\label{eq:henrard_hop}
\end{align}
Note that $\Pr \p{i \to j} + \Pr\p{i \to k} = 1$. This can be used directly in
conjunction with \autoref{se:area_ward} to understand for what initial
conditions each track can be observed and with what probabilities.
